---
title: "Climate tense analysis"
author: "Naia Morueta-Holme"
date: "January 20 2017"
output: html_document
---
```{r setup, include=F}
knitr::opts_chunk$set(comment = "#>")
```
### 1. Prepare data
####Clean workspace and load libraries
```{r message=FALSE}
#rm(list=ls())
require(likert)
require(ggplot2)
require(dplyr)
require(gmodels)
```

####Read in the raw data and subset
```{r, message=FALSE}
# NOTE: DOWNLOAD COPY FROM GOOGLE DRIVE AND CHANGE FILE NAME HERE ACCORDINGLY
rawData = read.csv('Data/Data_CC_tense_25Jan2017.csv', as.is=T) 

# Filter papers that have actually been checked
filledData = rawData[which(rawData$OK_for_analysis != ''),]

# Select papers that are OK for analysis, and the relevant columns
df = filledData[which(filledData$OK_for_analysis == 'yes'),
                    c('ID_old','Publication_year','Journal','paper_focus',
                      'tense','confidence_in_tense','climate_date_range',
                      'climate_date_min','climate_date_max',
                      'upd_climate_data_source','citation',
                      'geographic_extent_of_study','Data_collector_name')]

#lump continental and global
df$upd_geographic_extent = df$geographic_extent_of_study
df$upd_geographic_extent[which(df$geographic_extent_of_study=='continental' | df$geographic_extent_of_study=='global')] = 'continental/global'

#Add tense_lumpbad column
df$tense_lumpbad <- df$tense
df$tense_lumpbad[which(df$tense=='unclear' | df$tense=='inconsistent')] = 'ambiguous'

#Add tense_lump3 column for implicit/explicit
df$tense_lump3 <- df$tense_lumpbad
df$tense_lump3[which(df$tense=='past' | df$tense=='implicit past')] = 'any_past'
df$tense_lump3[which(df$tense=='present' | df$tense=='implicit present')] = 'any_present'

#Rename tense 'not temporal' to 'atemporal'
df$paper_focus[which(df$paper_focus=='not temporal')] = 'atemporal'

```

Table with impact factor for each journal (from www.citefactor.org)
```{r}
IF_lookup = data.frame(Journal = sort(unique(df$Journal)),IF2014=c(4.454, 5.469, 5.000, 13.042, 4.248, 5.694, 42.351, 9.809))
```

###2. Check the data

```{r}
#How many papers included in analysis?
table(filledData$OK_for_analysis)

#Is climate_data_range filled correctly? (only TRUE or FALSE valid)
table(df$climate_date_range)

#Do all papers have geographic extent information?
table(df$upd_geographic_extent)

#Do all papers have focus of study information?
table(df$paper_focus)

#Are all tense assignments confident?
table(df$confidence_in_tense)
```


Sampling per collector is not even, but OK
```{r echo=F}
#table(df$Data_collector_name)

TenseSum <-
  df %>%
  group_by(Data_collector_name) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

#TenseSum

TenseBasicPlot <- ggplot(TenseSum,aes(x=Data_collector_name,y=rel.freq)) +
  geom_bar(stat="identity") +
  coord_flip() +
  ggtitle("Papers included per data collector?") +
  ylab("% papers") +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.title.y = element_blank())

TenseBasicPlot
```

And journal representation is also uneven, so not analyzing differences
```{r echo=F}
TenseSum <-
  df %>%
  group_by(Journal) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

#TenseSum

TenseBasicPlot <- ggplot(TenseSum,aes(x=Journal,y=rel.freq)) +
  geom_bar(stat="identity") +
  coord_flip() +
  ggtitle("Papers included for each journal?") +
  ylab("% papers") +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.title.y = element_blank())

TenseBasicPlot
```

```{r, include=F, echo=F}
## Differences across time?
#Distribution of papers over time -> use to decide whether to work with proportions or raw counts
hist(rawData$Publication_year, breaks=seq(1979,2020,10), main="Papers per time period", ylim=c(0,250)) #all papers
hist(df$Publication_year, breaks=seq(1979,2020,10), add=T, col="grey")
legend(x="topright", legend=c("All", "In analysis"), fill=c("white", "grey"))


###But looks like there are more papers to be found!
hist(rawData$Publication_year, breaks=seq(1979,2020,5), main="Papers per time period", ylim=c(0,250)) #all papers
hist(df$Publication_year, breaks=seq(1979,2020,5), add=T, col="grey")
legend(x="topright", legend=c("All", "In analysis"), fill=c("white", "grey"))
```


### 3. Preliminary plots
```{r, include=T}
TenseSum <-
  df %>%
  group_by(tense) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

TenseSum

TenseBasicPlot <- ggplot(TenseSum,aes(x=tense,y=rel.freq)) +
  geom_bar(stat="identity") +
  coord_flip() +
  ggtitle("In what tense is climate referred to?") +
  ylab("% papers") +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.title.y = element_blank())

TenseBasicPlot
```
```{r include=T}
DateRangeSum <-
  df %>%
  group_by(climate_date_range) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))
DateRangeSum
```

```{r, include=F}
CitationSum <-
  df %>%
  group_by(citation) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

CitationSum
```


### 4. Crosstabulations for text (shown in figure)
```{r, include=F}
#Tables cross-classifying tense and groupings
## How to order the columns
order_cols <- c('past','implicit past', 'present', 'implicit present', 'ambiguous')

t1 <- CrossTable(df$paper_focus, df$tense_lumpbad)
res <-apply(t1$prop.row,2,function(x) {round(x*100,1)})

order_rows <- c('climate change effects','weather effects','atemporal')
ResFocus <- res[order_rows,order_cols]

t1 <- CrossTable(df$upd_geographic_extent, df$tense_lumpbad)
res <- apply(t1$prop.row,2,function(x) {round(x*100,1)})

order_rows <- c('continental/global','regional','site')
ResExtent <- res[order_rows,order_cols]

t1 <- CrossTable(df$climate_date_range, df$tense_lumpbad)
res <-apply(t1$prop.row,2,function(x) {round(x*100,1)})

order_rows <- c('TRUE','FALSE')
ResDateRange <- res[order_rows,order_cols]

t1 <- CrossTable(df$citation, df$tense_lumpbad)
res <-apply(t1$prop.row,2,function(x) {round(x*100,1)})

order_rows <- c('TRUE','FALSE')
ResCitation <- res[order_rows,order_cols]
```


```{r}
#Results based on geographic extent
ResExtent

#Results based on focus of paper
ResFocus

#Results based on whether date range is given
ResDateRange

#Results based on whether citation is given
ResCitation
```
<!-- Some more cross tabulations -->
<!-- ```{r, include=F} -->
<!-- t1 <- CrossTable(df$climate_date_range, df$citation) -->
<!-- res1 <-apply(t1$prop.row,2,function(x) {round(x*100,1)}) -->

<!-- t1 <- CrossTable(df$climate_data_source, df$upd_geographic_extent) -->
<!-- res2 <-apply(t1$prop.row,2,function(x) {round(x*100,1)}) -->


<!-- ``` -->


### 5. Final plot
Adapted from Matt's code
```{r echo=F}
### MANUALLY CONSTRUCT LIKERT BAR CHART
d <- df

# combine unclear and inconsistent tenses
# d$tns <- d$tense
# d$tns[grepl("inconsistent|unclear", d$tns)] <- "ambiguous"
# d$tns <- factor(d$tns, levels=rev(c("past", "implicit past", "ambiguous", "implicit present", "present")))
d$tns <- d$tense_lumpbad
d$tns <- factor(d$tns, levels=rev(c("past", "implicit past", "ambiguous", "implicit present", "present")))

# # clean invalid labels
# d$paper_focus[d$paper_focus=="***not temporal"] <- NA
# d$citation[!d$citation %in% c("TRUE", "FALSE")] <- NA
# d$geographic_extent_of_study[d$geographic_extent_of_study==""] <- NA

# gather factors
library(tidyr)
library(dplyr)
d <- gather(d, factor, level, paper_focus, upd_geographic_extent, citation, climate_date_range) %>%
      filter(!is.na(level)) %>%
      mutate(factor=factor(factor, 
                           levels=c("paper_focus", "upd_geographic_extent", "climate_date_range", "citation"),
                           labels=c("Study\nfocus", "Study\nextent", "Time\nperiod?", "Citation?")))

# p <- ggplot(d, aes(level, fill=tns)) + 
#       geom_bar(position="fill") +
#       scale_fill_manual(values=rev(c("darkred", "red", "gray80", "dodgerblue", "darkblue"))) +
#       coord_flip() +
#       facet_grid(factor~., scales="free", space="free") +
#       theme(panel.background=element_blank()) +
#       theme(axis.ticks.y=element_blank()) +
#       labs(fill="tense", y="proportion of papers", x=NULL)
# #ggsave("stacked_bar.png", width=7, height=5, units="in")


prop2likert <- function(x){
      y <- x
      for(i in 2:length(y)) y[i] <- y[i] + y[i-1] # top of each bar
      y0 <- c(0, y[1:4])
      ym <- (y + y0) / 2
      ym <- ym - ym[3]
      return(ym)
}

f <- d %>%
      
      # calculate proportions in each class
      group_by(factor, level, tns) %>%
      summarize(n=n()) %>%
      group_by(factor, level) %>%
      mutate(n=n/sum(n)) %>%
      
      # add 0's for non-observed tenses
      mutate(tns=as.character(tns), tns=sub(" ", "_", tns)) %>%
      spread(tns, n) %>%
      gather(tns, n, ambiguous:present) %>%
      mutate(n=ifelse(is.na(n), 0, n),
             tns=factor(tns, levels=c("past", "implicit_past", "ambiguous", 
                                      "implicit_present", "present"),
                        labels=c("past", "implicit past", "ambiguous", 
                                 "implicit present", "present"))) %>%
      
      # compute likert centers
      arrange(factor, level, tns) %>%
      group_by(factor, level) %>%
      mutate(x = prop2likert(n)) %>%
      
      # put levels in a sensical order
      ungroup() %>%
      mutate(level=factor(level, 
                          levels=c("atemporal", "weather effects", "climate change effects", 
                                   "site", "regional", "continental/global",
                                   "FALSE", "TRUE")))

p <- ggplot(f, aes(x=x*100, y=level, width=n*100, height=.9, fill=tns)) + 
      geom_tile() +
      facet_grid(factor~., scales="free", space="free") +
      scale_fill_manual(values=c("darkred", "red", "gray80", "dodgerblue", "darkblue")) +
      theme(panel.background=element_rect(fill="gray95"), panel.grid=element_blank(),
            axis.title.y=element_blank(), axis.ticks.y=element_blank(),
            legend.position="top") +
      scale_x_continuous(breaks=seq(-100, 100, 25), 
                         labels=abs(seq(-100, 100, 25)), 
                         limits=c(-100,100)) +
      labs(fill="Climate tense", x="percentage of papers")

p
#ggsave("likert_bars_v2.png", width=7, height=5, units="in")


```

