---
title: "Climate tense analysis"
author: "Naia Morueta-Holme"
date: "November 28 2016"
output: html_document
---
```{r setup, include=F}
knitr::opts_chunk$set(comment = "#>")
```
### Prepare data
Clean workspace and load libraries
```{r include=FALSE}
#rm(list=ls())
require(likert)
require(ggplot2)
require(dplyr)
require(gmodels)
```

Read in the raw data 
```{r}
# NOTE: DOWNLOAD COPY FROM GOOGLE DRIVE AND CHANGE FILE NAME HERE ACCORDINGLY
rawData = read.csv('Data/Data_CC_tense_01Oct2015_v2.csv', as.is=T) 
```

Clean the data
```{r message=F}
#Note: several of these clean-ups should be irrelevant once we have the full dataset
# Filter papers that have actually been checked
filledData = rawData[which(rawData$OK_for_analysis != ''),]
table(filledData$OK_for_analysis)

# Select papers that are OK for analysis, and the relevant columns
df = filledData[which(filledData$OK_for_analysis == 'yes'),
                    c('ID_old','Publication_year','Journal','paper_focus',
                      'tense','confidence_in_tense','climate_date_range',
                      'climate_date_min','climate_date_max',
                      'climate_data_source','citation',
                      'geographic_extent_of_study','Data_collector_name')]

# Filter papers where climate_data_range is filled correctly
table(df$climate_date_range)
df = subset(df, climate_date_range == T | df$climate_date_range == F)

# Filter papers with full geographic extent information (1 missing)
df = subset(df, geographic_extent_of_study != '')

# Filter papers where focus of study is filled correctly
table(df$paper_focus)
df = subset(df, paper_focus != "***not temporal")

```

Table with impact factor for each journal (from www.citefactor.org)
```{r}
IF_lookup = data.frame(Journal = sort(unique(df$Journal)),IF2014=c(4.454, 5.469, 5.000, 13.042, 4.248, 5.694, 42.351, 9.809))
```

Add tense_lumpbad column
```{r}
df$tense_lumpbad <- df$tense
df$tense_lumpbad[which(df$tense=='unclear' | df$tense=='inconsistent')] = 'unclear_inconsistent'

```

<!-- #-----------------# -->
<!-- # THE FUN BEGINS! # -->
<!-- #-----------------# -->
## Heads up!
Need to check a few tense assignments...
```{r echo=F}
table(df$confidence_in_tense)

TenseSum <-
  df %>%
  group_by(confidence_in_tense) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

#TenseSum

TenseBasicPlot <- ggplot(TenseSum,aes(x=confidence_in_tense,y=rel.freq)) +
  geom_bar(stat="identity") +
  coord_flip() +
  ggtitle("Confidence in tense assignment?") +
  ylab("% papers") +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.title.y = element_blank())
  
TenseBasicPlot
```

Sampling per collector is not even
```{r echo=F}
#table(df$Data_collector_name)

TenseSum <-
  df %>%
  group_by(Data_collector_name) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

#TenseSum

TenseBasicPlot <- ggplot(TenseSum,aes(x=Data_collector_name,y=rel.freq)) +
  geom_bar(stat="identity") +
  coord_flip() +
  ggtitle("Papers included per data collector?") +
  ylab("% papers") +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.title.y = element_blank())
  
TenseBasicPlot
```

And journal representation is also uneven
```{r echo=F}
#table(df$Journal)

TenseSum <-
  df %>%
  group_by(Journal) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

#TenseSum

TenseBasicPlot <- ggplot(TenseSum,aes(x=Journal,y=rel.freq)) +
  geom_bar(stat="identity") +
  coord_flip() +
  ggtitle("Papers included for each journal?") +
  ylab("% papers") +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.title.y = element_blank())
  
TenseBasicPlot
```

How about representation across time?
```{r echo=F}
## Differences across time?
#Distribution of papers over time -> use to decide whether to work with proportions or raw counts
hist(rawData$Publication_year, breaks=seq(1979,2020,10), main="Papers per time period", ylim=c(0,250)) #all papers
hist(df$Publication_year, breaks=seq(1979,2020,10), add=T, col="grey")
legend(x="topright", legend=c("All", "In analysis"), fill=c("white", "grey"))
```

But looks like there are more papers to be found!
```{r echo=F}
hist(rawData$Publication_year, breaks=seq(1979,2020,5), main="Papers per time period", ylim=c(0,250)) #all papers
hist(df$Publication_year, breaks=seq(1979,2020,5), add=T, col="grey")
legend(x="topright", legend=c("All", "In analysis"), fill=c("white", "grey"))
```


## Preliminary plots
```{r include=F}
TenseSum <-
  df %>%
  group_by(tense) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

TenseSum

TenseBasicPlot <- ggplot(TenseSum,aes(x=tense,y=rel.freq)) +
  geom_bar(stat="identity") +
  coord_flip() +
  ggtitle("In what tense is climate referred to?") +
  ylab("% papers") +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.title.y = element_blank())
  
TenseBasicPlot
```

```{r include=F}
#Adapt files for likert plots
d1 <-
  df %>%
  #filter(!is.na(Journal), !is.na(geographic_extent_of_study)) %>%
  select(tense)
#order the data
d1$tense <- factor(d1$tense,
                   c("past",
                     "implicit past",
                     "present",
                     "implicit present",
                     "inconsistent",
                     "unclear"))



#Lump inconsistent and unclear tenses
d2 <- 
  df %>%
  #filter(!is.na(Journal), !is.na(geographic_extent_of_study)) %>%
  select(tense_lumpbad)
#order the data
d2$tense_lumpbad <- factor(d2$tense_lumpbad,
                           c("past",
                             "implicit past",
                             "unclear_inconsistent",
                             "implicit present",
                             "present"))

# subset grouping variables
d1_groups <- 
  df %>%
  #filter(!is.na(Journal), !is.na(geographic_extent_of_study)) %>%
  select(Journal, geographic_extent_of_study, paper_focus)

d1_groups$geographic_extent_of_study <- factor(d1_groups$geographic_extent_of_study, 
                     c("site",
                       "regional",
                       "continental",
                       "global"))
d1_groups$Journal <- factor(d1_groups$Journal,
                            c("GCB",
                              "American_Naturalist",
                              "Ecology",
                              "Diversity_and_Distributions",
                              "Journal_of_Ecology",
                              "PNAS",
                              "Ecology_Letters",
                              "Nature"))

d1_groups$paper_focus <- factor(d1_groups$paper_focus,
                                c("not temporal",
                                  "weather effects",
                                  "climate change effects"))


d1_groups$decade <- sapply(df$Publication_year, function(x) {floor(x/10)*10})
d1_groups$decade <- factor(d1_groups$decade,
                           c("1980",
                             "1990",
                             "2000",
                             "2010"))

```


```{r echo=F}
likert_11 <- likert(d1, grouping = d1_groups$Journal)
title <- "How is climate referred to across journals?"
plot(likert_11, center=2.5) + ggtitle(title)

likert_12 <- likert(d1, grouping = d1_groups$geographic_extent_of_study)
title <- "Does tense depend on extent of study?"
plot(likert_12, center=2.5) + ggtitle(title)

likert_13 <- likert(d1, grouping = d1_groups$paper_focus)
title <- "Does tense depend on focus of study?"
plot(likert_13, center=2.5) + ggtitle(title)

likert_14 <- likert(d1, grouping = d1_groups$decade)
title <- "Does use of tense change over time?"
plot(likert_14, center=2.5) + ggtitle(title)
```

Plots lumping 'unclear' and 'inconsistent'
```{r echo=F}
likert_22 <- likert(d2, grouping = d1_groups$geographic_extent_of_study)
title <- "Does tense depend on extent of study?"
plot(likert_22) + ggtitle(title)

likert_23 <- likert(d2, grouping = d1_groups$paper_focus)
title <- "Does tense depend on focus of study?"
plot(likert_23) + ggtitle(title)

```


```{r include=F, echo=F}
#Tables cross-classifying tense and groupings (since plots look fishy...)
## How to order the columns
order_cols <- c('past','implicit past', 'present', 'implicit present', 'unclear_inconsistent')

t1 <- CrossTable(df$paper_focus, df$tense_lumpbad)
res <-apply(t1$prop.row,2,function(x) {round(x*100,1)})

order_rows <- c('climate change effects','weather effects','not temporal')
ResFocus <- res[order_rows,order_cols]

t1 <- CrossTable(df$geographic_extent_of_study, df$tense_lumpbad)
res <- apply(t1$prop.row,2,function(x) {round(x*100,1)})

order_rows <- c('global','continental','regional','site')
ResExtent <- res[order_rows,order_cols]

t1 <- CrossTable(df$climate_date_range, df$tense_lumpbad)
res <-apply(t1$prop.row,2,function(x) {round(x*100,1)})

order_rows <- c('TRUE','FALSE')
ResDateRange <- res[order_rows,order_cols]

t1 <- CrossTable(df$citation, df$tense_lumpbad)
res <-apply(t1$prop.row,2,function(x) {round(x*100,1)})

order_rows <- c('TRUE','FALSE')
ResCitation <- res[order_rows,order_cols]
```

Crosstabulations in percent
```{r}
ResFocus
ResExtent
ResDateRange
ResCitation
```
